---
title: "Internet Archive API Client"
author: "Lincoln Mullen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This API client for the [Internet Archive](https://archive.org/) is intended primarily for searching for items, retrieving metadata for items, and downloading the files associated with the items. The functions can be used with the pipe operator (`%>%`) from [magrittr](https://github.com/smbache/magrittr) and the data manipulation verbs in [dplyr](https://github.com/hadley/dplyr) to create pipelines from searching to downloading. For the full details of what is possible with the Internet Archive API, see their [advanced search help](https://archive.org/advancedsearch.php).

## Installation

Install the [development version](https://github.com/lmullen/internetarchive) from GitHub.

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("lmullen/internetarchive", build_vignettes = TRUE)
```

Then load the package. We will also use [dplyr](https://github.com/hadley/dplyr) for manipulating the retrieved data.

```{r}
library("internetarchive")
library("dplyr")
```

## Basic search and browse

The simplest way to search the Internet Archive is to use a keyword search. The following function searches for these keywords in the most important metadata fields, and returns a list of item identifiers.

```{r}
ia_keyword_search("isaac hecker")
```

You can pass an item identifier to the `ia_browse()` function to open an item in your browser. If you pass this function multiple identifiers, it will open only the first one.

```{r, eval=FALSE}
ia_browse("TheLifeOfFatherHecker")
```

## Advanced search

Usually it is more useful to perform an advanced search. You can construct an advanced search as a named character vector, where the names correspond to the fields. The following search, for instance, looks for items published by the American Tract Society in 1864. Run the function `ia_list_fields()` to see the list of accepted metadata fields. 

```{r}
ats_query <- c("publisher" = "american tract society", "year" = "1864")
ia_search(ats_query, num_results = 20)
```

You can change the number of items returned by the search using the `num_results =` argument, and you can request subsequent pages of results with the `page =` argument.

Notice that `ia_search()` and `ia_keyword_search()` both return a character vector of identifiers, so both can be used in the same way at the beginning of a pipeline.

## Getting item metadata and files

Once you have retrieved a list of items, you can retrieve their metadata and the list of files associated with the items.

To get a single item's metadata, you can pass its identifier to the `ia_get_items()` function.

```{r}
hecker <- ia_get_items("TheLifeOfFatherHecker")
```

The result is a list where the names of items in the list are the item identifiers, and the rest of the list is the metadata. This nested list can be difficult to work with, so the `ia_metadata()` returns a data frame of the metadata, and `ia_files()` returns a data frame of the files associated with the item.

```{r}
ia_metadata(hecker)
ia_files(hecker)
```

These functions can also retrieve the information for multiple items when used in a pipeline. Here we search for all the items about Hecker, retrieve their metadata, and turn it into a data frame. We then filter the data frame to get only the titles.

```{r}
ia_keyword_search("isaac hecker", num_results = 20) %>% 
  ia_get_items() %>% 
  ia_metadata() %>% 
  filter(field == "title") %>% 
  select(value)
```

## Downloading files

The `ia_download()` function will download all the files in a data frame returned from `ia_files()`. This function should be used with caution, and you should first filter the data frame to download only the files that you wish. In the following example, we retrieve a list of all the files associated with items published by the American Tract Society in 1864. Then we filter the list so we get only text files, then we pick only the first text file associated with each item. Finally we download the files to a directory we specify (in this case, a temporary directory).

```{r}
dir <- tempdir()
ia_search(ats_query) %>% 
  ia_get_items() %>% 
  ia_files() %>% 
  filter(type == "txt") %>% 
  group_by(id) %>% 
  slice(1) %>% 
  ia_download(dir = dir, overwrite = FALSE) %>% 
  glimpse()
```

Notice that `ia_download()` returns a modified version of the data frame that was passed to it, adding a column `local_file` with the path to the download files.

If the `overwrite =` argument is `FALSE`, then you can pass the same data frame of files to `ia_download()` and it will download only the files that it has not already downloaded.